Using Genetic Algorithms to Find Approximations for the Minimum Vertex Cover Problem

### 2018-02-23 paper.tex
stupid code folding erroe

% preferences
\graphicspath{ {images/} }
\newcommand{\lineSeparationLength}{2mm}
\hypersetup{colorlinks=true, allcolors=black}
\addbibresource{references.bib}
\ExecuteBibliographyOptions{
	giveninits=true,
	isbn=false,
	doi=false,
	url=false
}
\setlength{\parskip}{1em}
\newcommand{\textem}[1]{\textbf{\textit{#1}}}
\newtheorem{lemma}{Lemma}

@@

\begin{figure}[ht]
\centering
\includegraphics[width=1\textwidth]{ga}
\caption{general schema for genetic algorithms}
\label{fig:ga}
\end{figure}

@@

\subsection{Related Algorithms}
Since we are interested in covering all the edges of the
given graph by using as few nodes as possible, one might
be tempted to use a greedy-based heuristic to tackle the
mvcp. The algorithm consists in repeatedly selecting a
vertex of highest degree (the node that covers as many
of the remaining edges as possible), and removing all
of its incident edges. This is not a good strategy as
was demonstrated by Papadimitriou and Steiglitz ([10],
p. 407). \\
They considered regular graphs, each of which
consists of three levels. The first two levels have the same
number of nodes while the third level has two nodes less
than the number of nodes found on the previous two
levels. More precisely, each graph consists of $n = 3k + 4 (k \geq 1)$
nodes; $k+2$ nodes on the first level, labeled $1, 2, \dots, k+2$;
followed by $k+2$ nodes on the second level, labeled $k+3, \dots, 2k+4$;
while the $k$ nodes of the third level are labeled $2k+5, \dots, 3k+4$.
The regular graph for $k = 4$ can be found in figure~\ref{fig:g1}.
A minimum vertex cover is obtained by choosing all
the nodes of the second level, but the greedy strategy
would start with the nodes of the third level, since these
have highest degree. Consequently, the greedy strategy
finds a solution of size $2k + 2$ (since it has to select $k + 2$
nodes in addition to the nodes of the third level), while
the optimal solution has size $k + 2$.
However this algorithm is not totally useless. It may be shown that it always
achieves the ratio $\Theta(\lg n)$, where $n$ is the number of vertices.
That is, if $C^*(n)$ is the size of the optimal
solution and $C(n)$ is the size of the solution found by the aforementioned
algorithm. Then we have $\frac{C(n)}{C^*(n)} = \Theta(\lg n)$ [9], p-323.

@@

\begin{algorithm}
\caption{vercov}
\label{alg:vercov}
\begin{algorithmic}[1]

\State $C \gets \{ \}$
\Comment{$C$ contains the vertex cover being constructed}
\State $E' \gets E$
\While{$E' \neq \phi$}
\State randomly choose $ \langle u, v \rangle \in E'$
\State $C \gets C \cup \{u, v\}$
\State $E' \gets E' -
\{ \langle x, y \rangle \in E' : x=u \vee y=v \vee x=v \vee y=u \}$
\State \Comment{remove from $E'$ every edge incident on either $u$ or $v$}
\EndWhile
\State \Return $C$

@@

This is a \textit{2-approximation algorithm}. Which means that
the size of the vertex cover
it returns is guaranteed to be no more than twice the
size of an optimal vertex cover ([1], p. 968).
More precisely, if $C^*$ is the size of the optimal
solution and $C$ is the size of the solution returned by Algorithm~\ref{alg:vercov}
then the inequality $C \leq 2C^*$ must hold (notice that here the approximation
doesn't depend on the number of vertices $n$).

\subsection{GA for the mvcp}
In order to apply GA to solve the mvcp, a cover $V'$ is represented
by a binary vector $\vec{x} = x_1x_2 \dots x_n$ where $x_i = 1$ if
the $i^{\text{th}}$ node is in $V'$, and $x_i = 0$ if it is not.
Using this representation, genetic algorithm is applied to minimize the following
fitness function: [0] \\
\begin{equation}
\label{eq:f}
\begin{split}
f(\vec{x})	&= \sum_{i=1}^{n} \left( x_i+n \cdot (1-x_i) \cdot
			   \sum_{j=i}^{n} (1-x_j) e_{ij} \right) \\
 			&= \sum_{i=1}^{n} x_i + n \cdot \sum_{i=1}^{n}
			   \sum_{j=i}^{n} (1-x_i) \cdot (1-x_j) \cdot e_{ij}
\end{split}
\end{equation}
The term $ \sum_{i=1}^{n} x_i $ of $ f(\vec{x}) $ determines the size of the
potential vertex cover represented by $\vec{x}$, while the term
$ n \cdot \sum_{i=1}^{n} \sum_{j=i}^{n} (1-x_i) \cdot (1-x_j) \cdot e_{ij} $
penalizes sets $V'$ that
are not covers by adding a penalty of magnitude $n$ for
each edge $e_{ij}$ for which $i \notin V' \wedge j \notin V'$
(note that $e_{ii}$ is previously defined to be zero).
Consequently, any infeasible solution will have a fitness greater than
or equal to $n$, and for any feasible solution the second term will
drop to zero.
This fitness function was developed according to the
following design principles that are important for a
successful penalty function approach [11 $||$ 13 $||$ 7]:

@@

The experiments reported in section~\ref{sec:er} are performed
by using a genetic algorithm with a population size of
$u = 50$, a mutation rate $p_m = 1/n$, crossover rate
$p_c = 0.6$, proportional selection, and two-point crossover.
As reported in [2, 12], the latter is expected to perform
better than the traditional one-point crossover.

\section{Experimental Results}
\label{sec:er}
For the experimental tests, random graphs of size $n = 100$ with different
edge densities $d$, $d \in \{0.1, 0.2, 0.3, 0.4, 0.5\}$, are used.
an edge density of d = 0:1 means that an edge is placed between two
nodes with a probability of 0.1.

>>>>

% preferences
\graphicspath{ {images/} }
\newcommand{\lineSeparationLength}{2mm}
\hypersetup{colorlinks=true, allcolors=black}
\addbibresource{references.bib}
\ExecuteBibliographyOptions{
	giveninits=true,
	isbn=false,
	doi=false,
	url=false
}
\setlength{\parskip}{1em}
\newcommand{\textem}[1]{\textbf{\textit{#1}}}
\newtheorem{lemma}{Lemma}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\algnewcommand\algorithmicto{\textbf{to}}
\algrenewtext{For}[3]%
{\algorithmicfor\ #1 = #2 \algorithmicto\ #3 \algorithmicdo}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\textwidth]{ga}
\caption{general schema for genetic algorithms}
\label{fig:ga}
\end{figure}

@@

\subsection{Related Algorithms}
\label{sec:ra}
Since we are interested in covering all the edges of the
given graph by using as few nodes as possible, one might
be tempted to use a greedy-based heuristic to tackle the
mvcp. The algorithm consists in repeatedly selecting a
vertex of highest degree (the node that covers as many
of the remaining edges as possible), and removing all
of its incident edges. This is not a good strategy as
was demonstrated by Papadimitriou and Steiglitz ([10],
p. 407). \\
They considered regular graphs, each of which
consists of three levels. The first two levels have the same
number of nodes while the third level has two nodes less
than the number of nodes found on the previous two
levels. More precisely, each graph consists of $n = 3k + 4 (k \geq 1)$
nodes; $k+2$ nodes on the first level, labeled $1, 2, \dots, k+2$;
followed by $k+2$ nodes on the second level, labeled $k+3, \dots, 2k+4$;
while the $k$ nodes of the third level are labeled $2k+5, \dots, 3k+4$.
The regular graph for $k = 3$ can be found in figure~\ref{fig:ps}.
A minimum vertex cover is obtained by choosing all
the nodes of the second level, but the greedy strategy
would start with the nodes of the third level, since these
have highest degree. Consequently, the greedy strategy
finds a solution of size $2k + 2$ (since it has to select $k + 2$
nodes in addition to the nodes of the third level), while
the optimal solution has size $k + 2$.
However this algorithm is not totally useless. It may be shown that it always
achieves the ratio $\Theta(\lg n)$\footnote{$f(n) = \Theta(g(n)) \iff
\exists \alpha, \beta, n_0 : \forall n \geq n_0, \ 
\alpha \cdot g(n) \leq f(n) \leq \beta \cdot g(n)
$},
where $n$ is the number of vertices.
That is, if $C^*(n)$ is the size of the optimal
solution and $C(n)$ is the size of the solution found by the aforementioned
algorithm. Then we have $\frac{C(n)}{C^*(n)} = \Theta(\lg n)$ [9], p-323.

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\textwidth]{ps}
\caption{Construction of the regular graph after Papadimitriou
and Steiglitz with $k=3$.}
\label{fig:ps}
\end{figure}

@@

\begin{algorithm}
\caption{vercov($G = (V, E)$)}
\label{alg:vercov}
\begin{algorithmic}[1]

\State $C \gets \{ \}$
\Comment{$C$ contains the vertex cover being constructed}
\State $E' \gets E$
\While{$E' \neq \phi$}
\State randomly choose $ \langle u, v \rangle \in E'$
\label{alg:rand}
\State $C \gets C \cup \{u, v\}$
\State $E' \gets E' -
\{ \langle x, y \rangle \in E' : x=u \vee y=v \vee x=v \vee y=u \}$
\State \Comment{remove from $E'$ every edge incident on either $u$ or $v$}
\EndWhile
\State \Return $C$

@@

This is a \textit{2-approximation algorithm}. Which means that
the size of the vertex cover
it returns is guaranteed to be no more than twice the
size of an optimal vertex cover ([1], p. 968).
More precisely, if $C^*$ is the size of the optimal
solution and $C$ is the size of the solution returned by algorithm~\ref{alg:vercov}
then the inequality $C \leq 2C^*$ must hold (notice that here the approximation
doesn't depend on the number of vertices $n$).

\subsection{GA for the mvcp}
\label{sec:GAmvcp}
In order to apply GA to solve the mvcp, a cover $V'$ is represented
by a binary vector $\vec{x} = x_1x_2 \dots x_n$ where $x_i = 1$ if
the $i^{\text{th}}$ node is in $V'$, and $x_i = 0$ if it is not.
Using this representation, genetic algorithm is applied to minimize the following
fitness function: [0] \\
\begin{equation}
\label{eq:f}
\begin{split}
f(\vec{x})	&= \sum_{i=1}^{n} \left( x_i+n \cdot (1-x_i) \cdot
			   \sum_{j=i}^{n} (1-x_j) e_{ij} \right) \\
 			&= \sum_{i=1}^{n} x_i + n \cdot \sum_{i=1}^{n}
			   \sum_{j=i}^{n} (1-x_i) \cdot (1-x_j) \cdot e_{ij}
\end{split}
\end{equation}
The term $ \sum_{i=1}^{n} x_i $ of $ f(\vec{x}) $ determines the size of the
potential vertex cover represented by $\vec{x}$, while the term
$ n \cdot \sum_{i=1}^{n} \sum_{j=i}^{n} (1-x_i) \cdot (1-x_j) \cdot e_{ij} $
penalizes sets $V'$ that
are not covers by adding a penalty of magnitude $n$ for
each edge $e_{ij}$ for which $i \notin V' \wedge j \notin V'$
(note that $e_{ii}$ is previously defined to be zero).
Consequently, any infeasible solution will have a fitness greater than
or equal to $n$, and for any feasible solution the second term will
drop to zero.
This fitness function was developed according to the
following design principles that are important for a
successful penalty function approach [11 $||$ 13 $||$ 7]:

@@

The experiments reported in section~\ref{sec:er} are performed
by using a genetic algorithm with a population size of
$u = 50$, a mutation rate $p_m = 1/n$, crossover rate
$p_c = 0.6$, proportional selection, and two-point crossover.
As reported in [2, 12], the latter is expected to perform
better than the traditional one-point crossover.
The termination criterion is specified in section~\ref{sec:er}.

\section{Experimental Results}
\label{sec:er}
For the experimental tests,
the genetic algorithm software package GENEsYs is used [1]misp.
This implementation is based on the widely used GENESIS software by Grefenstette
(see [3], pp. 374-377 misp), but allows for more 
flexibility concerning genetic operators and data monitoring.
The parameter settings for our experiments are
given in section~\ref{sec:GAmvcp}.

Multiple problem instances are used to demonstrate the applicability of GAs to tackle the mvcp
The results in table~\ref{fig:t1}
and table~\ref{fig:t2} correspond to the first test set, which consists of
five random graphs of size $n = 100$ with different
edge densities $d$, $d \in \{0.1, 0.2, 0.3, 0.4, 0.5\}$, are used.
An edge density of $d = 0.1$ means that an edge is placed between two
nodes with a probability of 0.1.
Moreover we construct the graphs in a way that guarantees an optimum of quality 55
(we set $k = 55$ in algorithm~\ref{alg:gen}).
In addition to the regular graphs introduced by
Papadimitriou and Steiglitz and described in section~\ref{sec:ra}.
Recall that these graphs contain $n=3k+4 (k \geq 1)$ nodes distributed on three levels.
They can be scaled up by choosing high values for $k$.
A graph instance of the regular graph of size $n = 100 (k=32)$ is chosen.

\State randomly select $V' = \{i_1, \dots, i_k\} \subseteq V = \{1, \dots, n\}$
\Comment{randomly select k different nodes}
\For

\State $C \gets \{ \}$
\Comment{$C$ contains the vertex cover being constructed}
\State $E' \gets E$
\While{$E' \neq \phi$}
\State randomly choose $ \langle u, v \rangle \in E'$
\label{alg:rand}
\State $C \gets C \cup \{u, v\}$
\State $E' \gets E' -
\{ \langle x, y \rangle \in E' : x=u \vee y=v \vee x=v \vee y=u \}$
\State \Comment{remove from $E'$ every edge incident on either $u$ or $v$}
\EndWhile
\State \Return $C$

\end{algorithmic}
\end{algorit
For each of these graphs, a total of
$N = 100$ independent runs of the genetic algorithm is performed.
The results are summarized in table~\ref{fig:t1} for the best
fitness values that were encountered during the 100 runs.
For each problem instance, the different fitness values that were
obtained and their frequencies are recorded.
Furthermore, the average fitness value $\bar{f}$ over
all 100 runs is indicated at the bottom of the table.
For example, the first column in table~\ref{fig:t1} indicates
that the best obtained fitness value in 1 of the 100 runs was $f(\vec{x}) = 53$,
the best obtained fitness value in 1 of the 100 runs was $f(\vec{x}) = 54$,
the best obtained fitness value in 3 of the 100 runs was $f(\vec{x}) = 55$, and so on...
And the average of these values is indicated by $\bar{f}$, which is the average value of
the best obtained fitness values over the 100 runs.

As a termination criterion for the genetic algorithm, a bound for the total number of function evaluations is set.
For example The total number of function evaluations per single run is chosen to be $2 \cdot 10^4$ in table~\ref{fig:t1}.
That is, the algorithm terminates when $2 \cdot 10^4$ evaluations of the fitness function
had occurred.
This bound is indicated as an index $t$ in the notation $f_t(\vec{x})$.
Consequently, only a small fraction of the search space is explored by the genetic algorithm.
This fraction is about $ 2 \cdot 10^4 / 2^{100} \approx 1.6 \cdot 10^{-24} \% $ for the first test set.
and it is about $ 4 \cdot 10^4 / 2^{200} \approx 2.5 \cdot 10^{-54} \% $ for the second test set.

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\textwidth]{t1}
\caption{general schema for genetic algorithms}
\label{fig:t1}
\end{figure}

For each of the problems in the first test set, a total of $N=100$
independent runs of the vercov algorithm is performed. The results are summarized in table~\ref{fig:t2}.
Notice that for one problem instance, the results may differ in each run due to the
non-determinism of the algorithm (see algorithm~\ref{alg:vercov} line~\ref{alg:rand}).

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\textwidth]{t2}
\caption{general schema for genetic algorithms}
\label{fig:t2}
\end{figure}

To test the behaviour of the genetic algorithm
as well as the vercov algorithm for an even
larger problem size, The same experiments were performed for larger graphs
The results in table~\ref{fig:t3}
and table~\ref{fig:t4} correspond to the second test set, which consists of
five random graphs of size $n = 200$ with different
edge densities $d$, $d \in \{0.1, 0.2, 0.3, 0.4, 0.5\}$, are used.
Moreover we construct the graphs in a way that guarantees an optimum of quality 110
(see algorithm~\ref{alg:gen}).
In addition to a graph instance of the regular graphs introduced by
Papadimitriou and Steiglitzof of size $n = 220 (k=66)$.
In this case, the genetic algorithm
was allowed to run for $4 \cdot 10^4$ function evaluations.
The corresponding results obtained by the
genetic algorithm are shown in table~\ref{fig:t3}, while the results
from the vercov algorithm are presented in table~\ref{fig:t4}.
See the previous paragraphs for a better understanding of these results.

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\textwidth]{t3}
\caption{general schema for genetic algorithms}
\label{fig:t3}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\textwidth]{t4}
\caption{general schema for genetic algorithms}
\label{fig:t4}
\end{figure}

\subsection{expermental observations}
